# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
data=read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
### make a histogram of total number of steps taken per day
```{r}
# split by date
dateGroup=split(data,data$date)
stepsPerDay=sapply(dateGroup,function(ele) sum(data.frame(ele)$steps,na.rm=T))
hist(stepsPerDay,breaks=20)
```
### Calculate and report the mean and median total number of steps taken per day

```{r}
meanSteps=mean(stepsPerDay)
medianSteps=median(stepsPerDay)
```

The mean and median total number of steps taken per day are `r meanSteps` and `r medianSteps` respecitively.

## What is the average daily activity pattern?
### Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
# calculate the mean steps for each interval
intervalGroup=split(data,data$interval)
stepsPerInterval=sapply(intervalGroup,function(ele) mean(data.frame(ele)$steps,na.rm=T))
plot(names(stepsPerInterval),stepsPerInterval,type="l",xlab="5-min interval",ylab="steps")
```
### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
index=which(stepsPerInterval==max(stepsPerInterval,na.rm=T))
intervalName=names(stepsPerInterval)[index]
```
The `r intervalName` interval has the maximum number of steps on average.
## Imputing missing values

### Calculate and report the total number of missing values in the dataset 
```{r}
narows=dim(data)[1]-sum(complete.cases(data))
```

`r narows` observations in the original data have missing values.

### Using the means of 5-minute intervals to fill in all of the missing values in the dataset.

```{r}
# calculate the mean steps for each interval
intervalGroup=split(data,data$interval)
stepsPerInterval=sapply(intervalGroup,function(ele) mean(data.frame(ele)$steps,na.rm=T))
# use the mean steps to fill NAs
naRows=!complete.cases(data)
indexes=which(naRows)
# make a copy of the data
assign('data3',data)
for(i in 1:length(indexes))
{
    data3[indexes[i],1]=stepsPerInterval[[as.character(data[indexes[i],3])]]
}
```

### The above codes also create a new dataset called data3 that is equal to the original dataset but with the missing data filled in.

### Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
# split by date
dateGroup2=split(data3,data$date)
stepsPerDay2=sapply(dateGroup2,function(ele) sum(data.frame(ele)$steps,na.rm=T))
hist(stepsPerDay2,breaks=20)

meanSteps2=mean(stepsPerDay2)
medianSteps2=median(stepsPerDay2)
```

The new mean and median total number of steps taken per day are `r meanSteps2` and `r medianSteps2` respecitively.

These two new values are different from the mean and median calculated in part 1. 

Imputing missing data resulted in larger mean and median values.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
days=function(string)
{
    day=weekdays(as.POSIXlt(string))
    if(day!="Saturday" & day!="Sunday")
    {
        return("weekday")
    }
    else
    {
        return("weekend")
    }
}

day=rep(NA,dim(data)[1])
for(i in 1:dim(data)[1])
{
    day[i]=days(data$date[i])    
}

data4=transform(data3,wday=day)

weekday=subset(data4,wday=="weekday")
weekend=subset(data4,wday=="weekend")

par(mfrow=c(2,1))
# calculate the mean steps for each interval
intervalGroup3=split(weekend,weekend$interval)
stepsPerInterval3=sapply(intervalGroup3,function(ele) mean(data.frame(ele)$steps,na.rm=T))
plot(names(stepsPerInterval3),stepsPerInterval3,type="l",xlab="5-min interval",ylab="steps",main="weekend")

# calculate the mean steps for each interval
intervalGroup4=split(weekday,weekday$interval)
stepsPerInterval4=sapply(intervalGroup4,function(ele) mean(data.frame(ele)$steps,na.rm=T))
plot(names(stepsPerInterval4),stepsPerInterval4,type="l",xlab="5-min interval",ylab="steps",main="weekday")

par(mfrow=c(1,1))
```

